The function simulates outcomes for decision options. The `simulate_outcomes_for_decisions` function is a Monte Carlo simulation framework designed to evaluate multiple decision options against predefined outcomes, incorporating stochastic variability (variability_list) to reflect real-world uncertainties. This function is intended for use in scenarios where decisions have probabilistic impacts on outcomes, such as ecological management, conservation planning, or policy evaluation. Source the  `simulate_outcomes_for_decisions` function with `source`.

```{r source_simulate_outcomes_for_decisions}
source("functions/simulate_outcomes_for_decisions.R")
```

### Function Purpose

The function calculates the expected outcomes for a series of decision options under uncertainty. Each decision option is evaluated against a set of outcomes using a weighted linear model, with added random variability_list to account for variability. The outputs are stored in a structured format for further analysis.

Function Inputs include: 

`decision_options`, a list where each element is a named vector representing a specific decision option. Each vector contains binary values (1 or 0), indicating whether a particular decision variable is active or inactive. For example: `control_access = c(1, 0, 0, 0, 0)` indicates that `control_access` is the first option in the list of decision options. 

`outcome_effects`, a matrix where rows represent outcomes (e.g., conservation success, financial cost). The columns correspond to decision variables. The elements specify the weight or influence of a decision variable on a given outcome. i.e. `matrix(c(0.8, 0.6, 0.5, 0.6, 0.7),nrow = 1)` for eh species Conservation outcome_effects corresponding to the columns of `decision_options`. 

The `variability_list` is a named list of functions, with each function generating random noise specific to a specific outcome. This allows for custom noise distributions to reflect expert knowledge or empirical data on variability. i.e. `list(SpeciesConservation = function(n) rnorm(n, mean = 0, sd = 0.05))` is a function assigning the noise to the interaction between species conservation and the decision options. 

`n_simulations` is the number of Monte Carlo iterations to perform for each combination of decision and outcome. 

The function returns `results` as a nested list with: Decision options (e.g., `control_access`). Outcome names (e.g., `SpeciesConservation`). Values: A vector of simulated outcome values for each combination of decision and outcome.

## Simulation Workflow

### Decision Options:

The function iterates over each decision option in `decision_options`. For each decision option, the function calculates all outcomes defined in the `outcome_effects` matrix. Outcomes are calculated as a weighted sum of decision variables (defined by the `outcome_effects` matrix) plus a random variability_list term. `variability_list` is generated, allowing for outcome-specific variability. Results are stored in a nested list, enabling easy access for further analysis.

For a conservation project, this function will be used to simulate the effects of different park management strategies (e.g., restricting access, providing seedlings) on:

- Species conservation: Probability of preserving species.
- Dietary diversity: Impact on local food systems.
- Expenditure: Financial cost to the park service.
- Cultural heritage: Loss of traditional knowledge.

By comparing the simulated outcomes for each strategy, decision-makers can identify trade-offs and prioritize actions. The function can handle multiple decision options and outcomes. It incorporates expert-defined `variability_list` distributions for outcome variability. It also explicitly separates decision logic, outcome effects, and `variability_list`, ensuring clarity and reproducibility. For each decision option and outcome, the function calculates the **Outcome Value** using the following formula:

\[
\text{Outcome Value} = \sum_{i=1}^{n} \text{Decision}_i \times \text{Weight}_i + \text{variability_list}
\]

Where

- \(\text{Decision}_i\): A binary value (\(1\) or \(0\)) indicating if the \(i\)-th decision is active.
- \(\text{Weight}_i\): The influence of the \(i\)-th decision variable on the outcome.
- \(\text{variability_list}\): A random term representing variability or uncertainty, specific to the outcome and decision, as defined by the `variability_list`.

```{r forest_access_graph}
# Load necessary libraries
library(DiagrammeR)
library(DiagrammeRsvg)
library(rsvg)

# Create the updated graph
forest_access_graph <- grViz("
  digraph forest_access_model {
    graph [layout = dot, rankdir = LR, splines = line]

    // Define node styles
    node [shape = box, fontname = Helvetica]

    // Define nodes
    ControlAccess [label = 'Control Access']
    ProvideAccessControlHarvest [label = 'Provide Access + Control Harvest']
    AllowAccessForestParts [label = 'Allow Access to Forest Parts']
    AllowAccessLimitedTime [label = 'Allow Access for Limited Time']
    ProvideSeedlings [label = 'Provide Seedlings']

    SpeciesConservation [label = 'Species Conservation']
    DietaryDiversity [label = 'Dietary Diversity']
    SustainableFoodEnv [label = 'Sustainable Food Environment']
    ParkExpenditure [label = 'Park Expenditure']
    TraditionalKnowledge [label = 'Traditional Knowledge']

    // Define relationships between nodes
    ControlAccess -> {SpeciesConservation DietaryDiversity SustainableFoodEnv ParkExpenditure TraditionalKnowledge}
    ProvideAccessControlHarvest -> {SpeciesConservation DietaryDiversity SustainableFoodEnv ParkExpenditure TraditionalKnowledge}
    AllowAccessForestParts -> {SpeciesConservation DietaryDiversity SustainableFoodEnv ParkExpenditure TraditionalKnowledge}
    AllowAccessLimitedTime -> {SpeciesConservation DietaryDiversity SustainableFoodEnv ParkExpenditure TraditionalKnowledge}
    ProvideSeedlings -> {SpeciesConservation DietaryDiversity SustainableFoodEnv ParkExpenditure TraditionalKnowledge}
  }
")

forest_access_graph
```
In this model ParkExpenditure refers to the additional costs incurred (or saved) by implementing food access strategies, more is worse. All other values are positive, more is better.
