
## Add species with simplified initial input

Use the `add_species_data.R` function with other functions from the `tidyverse`

```{r species_params}
source("functions/add_species_data.R")
library(tidyverse)
```

The `Effect` represents the magnitude and direction of influence that a specific decision option has on an outcome for a given species. Positive values indicate a beneficial effect and negative values visa versa. For instance, a value of 0.8 for `control_access` on `Species Conservation` implies a strong positive effect, whereas -0.2 for `provide_seedlings` on `Dietary Diversity` indicates a slight reduction in dietary diversity. A value of 0 means no effect. The Effect value represents the proportional change in the outcome relative to its current state for a given species. The measure allows for a direct comparison of how different interventions influence species-level outcomes.

The SD (also `Variability_SD`) quantifies the uncertainty or variability in the effect values. A smaller SD (e.g., 0.05) indicates high confidence in `Effect`, suggesting that the impact is consistent across scenarios. Larger SD (e.g., 0.2) reflects greater uncertainty, for example, an SD of 0.05 for `control_access` on `Species Conservation` indicates low variability and high confidence, whereas an SD of 0.15 for `provide_seedlings` on `Dietary Diversity` suggests moderate uncertainty. 

## Decision options for sustainable forest food

Define `decision_options` for managing forest access (limited time, area or volume of harvest) or species access (seedlings provided)

```{r define_options_outcomes}
# Define decision options and outcomes
decision_options <- c("control_access", "provide_access_control_harvest", 
                      "allow_access_forest_parts", "allow_access_limited_time", "provide_seedlings")

# Define readable labels
decision_labels <- c(
  "control_access" = "Control Access",
  "provide_access_control_harvest" = "Regulated Harvesting",
  "allow_access_forest_parts" = "Forest Access",
  "allow_access_limited_time" = "Limited Access",
  "provide_seedlings" = "Seedling Provision"
)

```

## Outcomes of sustainable forest food decisions

Define decision `outcomes` for diets and conservation

```{r show_outcomes}
outcomes <- c("SpeciesConservation", "DietaryDiversity", "SustainableFoodEnv", 
              "ParkExpenditure", "TraditionalKnowledge")

outcome_labels <- c(
  "SpeciesConservation" = "Species Conservation",
  "DietaryDiversity" = "Dietary Diversity",
  "SustainableFoodEnv" = "Sustainable Food Environment",
  "ParkExpenditure" = "Park Expenditure",
  "TraditionalKnowledge" = "Traditional Knowledge"
)

```

### Species Classification

Below is the list of species mentioned and observed as useful in the field. The list is organized by functional role of species in conservation, food security. Although we are mainly concerned about forest conservation in protected forests we also include information about the role of the species in agroforestry systems since many of the species are kept in local homegardens by those community members who have access and land.


- **Mushrooms**:  
  - *Auricularia heimuer* (Jelly ear fungus)  
  - *Volvariella volvacea* (Paddy straw mushroom)  

- **Grasses & Bamboo**:  
  - *Ampelocalamus patellaris* (Bamboo)  
  - *Arundinaria baviensis* (Bamboo)  
  - *Bambusa blumeana* (Thorny bamboo)  
  - *Coix lacryma-jobi* (Job’s tears, grain)  
  - *Dendrocalamus velutinus* (Bamboo)  

- **Ferns**:  
  - *Diplazium esculentum* (Edible fern)  
  - *Nephrolepis cordifolia* (Sword fern)  

- **Shrubs & Small Trees**:  
  - *Bischofia javanica* (Bishop wood, tree)  
  - *Canarium bengalense* (Canarium tree)  
  - *Canarium nigrum* (Black canarium tree)  
  - *Canarium tramvangum* (Canarium tree)  
  - *Euodia bodinieri* (Rutaceae tree/shrub)  
  - *Maesa perlarius* (Shrub)  
  - *Schefflera heptaphylla* (Umbrella tree)  

- **Climbers & Vines**:  
  - *Cissus hastata* (Vine)  
  - *Erythropalum scandens* (Climber)  
  - *Hodgsonia macrocarpa* (Climber, liana)  
  - *Paederia lanuginosa* (Climber)  
  - *Passiflora foetida* (Passionflower vine)  

- **Herbs & Vegetable Crops**:  
  - *Alpinia xanthioides* (Ginger family)  
  - *Begonia aptera* (Begonia, herb)  
  - *Begonia tonkinensis* (Begonia, herb)  
  - *Bidens pilosa* (Blackjack, medicinal herb)  
  - *Centella asiatica* (Gotu kola, medicinal herb)  
  - *Costus speciosus* (Spiral ginger, herb)  
  - *Crassocephalum crepidioides* (Edible herb)  
  - *Curculigo orchioides* (Orchioides, herb)  
  - *Dioscorea hamiltonii* (Yam, tuber crop)  
  - *Gynura procumbens* (Longevity spinach, herb)  
  - *Gynura sarmentosa* (Herb)  
  - *Melientha suavis* (Pak wan, vegetable crop)  
  - *Momordica cochinchinensis* (Gac fruit, climbing vegetable)  
  - *Peperomia pellucida* (Shiny bush, herb)  
  - *Physalis angulata* (Ground cherry, herbaceous plant)  
  - *Piper lolot* (Wild betel leaf, herbaceous vine)  
  - *Plantago major* (Plantain herb)  

- **Trees**:  
  - *Baccaurea ramiflora* (Burmese grape tree)  
  - *Ficus auriculata* (Fig tree)  
  - *Ficus callosa* (Fig tree)  
  - *Ficus heterophyllus* (Fig tree)  
  - *Ficus racemosa* (Cluster fig tree)  
  - *Nephelium lappaceum* (Rambutan tree)  
  - *Oroxylum indicum* (Tree)  
  - *Rubus cochinchinensis* (Tree-like shrub)  
  - *Saurauia tristyla* (Tree)  
  - *Zanthoxylum rhetsa* (Prickly ash tree)  

- **Flowering Plants (Ornamentals, Medicinal, & Wild Edibles)**:  
  - *Clausena indica* (Wild edible, citrus family)  
  - *Clerodendrum chinense* (Glory bower, flowering plant)  
  - *Clerodendrum cyrtophyllum* (Wild medicinal plant)  
  - *Clerodendrum kaempferi* (Ornamental plant)  
  - *Clerodendrum philippinum* (Glory flower, ornamental)  
  - *Clerodendrum quadriloculare* (Ornamental, medicinal)  
  - *Wedelia trilobata* (Creeping daisy, ground cover)  
  - *Polygonum perfoliatum* (Mile-a-minute vine, invasive)  


## Create species files

Create `species_files` a list of species and their corresponding `.R` files from all field visits. `species_files` is used for updating in the section Update based on field observations @update_model

```{r species_files_first_visits}
# List of species and their corresponding `.R` files
species_files <- list(
  # === One very common species 
  bidens_pilosa = "data/effects_sd_map_bidens_pilosa.R",               
  # Bidens pilosa
  
  # === Learning in early 2024
  
  amomum_villosum = "data/effects_sd_map_amomum_villosum.R",           
  # Amomum villosum
  begonia = "data/effects_sd_map_begonia.R",                 
  # Begonia
  bischofia_javanica = "data/effects_sd_map_bischofia_javanica.R",    
  # Bischofia javanica Blum
  cheilocostus_speciosus = "data/effects_sd_map_cheilocostus_speciosus.R", 
  # Cheilocostus speciosus
  cyclosorus_acuminatus = "data/effects_sd_map_cyclosorus_acuminatus.R",
  # Cyclosorus acuminatus (Houtt)
  erythropalum_scandens = "data/effects_sd_map_erythropalum_scandens.R", 
  # Erythropalum scandens Blume
  ficus_auriculata = "data/effects_sd_map_ficus_auriculata.R", 
  # Ficus auriculata
  ficus_callosa = "data/effects_sd_map_ficus_callosa.R",     
  # Ficus callosa Willd.
  gynura_procumbens = "data/effects_sd_map_gynura_procumbens.R",        
  # Gynura procumbens (Lour.)
  melientha_suavis = "data/effects_sd_map_melientha_suavis.R",      
  # Melientha suavis
  musa_brachycarpa = "data/effects_sd_map_musa_brachycarpa.R",           
  # Musa brachycarpa Back
  ophiopogon_reptans = "data/effects_sd_map_ophiopogon_reptans.R",   
  # Ophiopogon reptans
  rubus_cochinchinensis = "data/effects_sd_map_rubus_cochinchinensis.R",     
  # Rubus cochinchinensis
  saurauia_tristyla = "data/effects_sd_map_saurauia_tristyla.R",      
  # Saurauia tristyla DC.
  schefflera_heptaphylla = "data/effects_sd_map_schefflera_heptaphylla.R", 
  # Schefflera heptaphylla
  sterculia_foetida = "data/effects_sd_map_sterculia_foetida.R",     
  # Sterculia foetida L.
  tetrapanax_papyriferus = "data/effects_sd_map_tetrapanax_papyriferus.R", 
  # Tetrapanax papyriferus
  tradescantia_pallida = "data/effects_sd_map_tradescantia_pallida.R", 
  # Tradescantia pallida
    clerodendrum_cyrtophyllum = "data/effects_sd_map_clerodendrum_cyrtophyllum.R", 
  
  # === From later visits Oct 2024
  
  # Clerodendrum cyrtophyllum
  crassocephalum_crepidioides = "data/effects_sd_map_crassocephalum_crepidioides.R", 
  # Crassocephalum crepidioides
  centella_asiatica = "data/effects_sd_map_centella_asiatica.R", 
  # Centella asiatica
  dioscorea_hamiltonii = "data/effects_sd_map_dioscorea_hamiltonii.R", 
  # Dioscorea hamiltonii
  artemisia_vulgaris = "data/effects_sd_map_artemisia_vulgaris.R", 
  # Artemisia vulgaris
  
  # === From later visits after Jan 2025
  
  gynura_sarmentosa = "data/effects_sd_map_gynura_sarmentosa.R", 
  # Gynura sarmentosa
  euodia_bodinieri = "data/effects_sd_map_euodia_bodinieri.R", 
  # Euodia bodinieri
  nephelium_lappaceum = "data/effects_sd_map_nephelium_lappaceum.R", 
  # Nephelium lappaceum
  hodgsonia_macrocarpa = "data/effects_sd_map_hodgsonia_macrocarpa.R", 
  # Hodgsonia macrocarpa
  baccaurea_ramiflora = "data/effects_sd_map_baccaurea_ramiflora.R", 
  # Baccaurea ramiflora
  cissus_hastata = "data/effects_sd_map_cissus_hastata.R", 
  # Cissus hastata
  maesa_perlarius = "data/effects_sd_map_maesa_perlarius.R", 
  # Maesa perlarius
  nephrolepis_cordifolia = "data/effects_sd_map_nephrolepis_cordifolia.R", 
  # Nephrolepis cordifolia
  piper_lolot = "data/effects_sd_map_piper_lolot.R", 
  # Piper lolot
  passiflora_foetida = "data/effects_sd_map_passiflora_foetida.R", 
  # Passiflora foetida
  plantago_major = "data/effects_sd_map_plantago_major.R", 
  # Plantago major
  zanthoxylum_rhetsa = "data/effects_sd_map_zanthoxylum_rhetsa.R", 
  # Zanthoxylum rhetsa
  costus_speciosus = "data/effects_sd_map_costus_speciosus.R", 
  # Costus speciosus
  clerodendrum_chinense = "data/effects_sd_map_clerodendrum_chinense.R", 
  # Clerodendrum chinense
  paederia_lanuginosa = "data/effects_sd_map_paederia_lanuginosa.R", 
  # Paederia lanuginosa
  clausena_indica = "data/effects_sd_map_clausena_indica.R", 
  # Clausena indica
  auricularia_heimuer = "data/effects_sd_map_auricularia_heimuer.R", 
  # Auricularia heimuer
  volvariella_volvacea = "data/effects_sd_map_volvariella_volvacea.R", 
  # Volvariella volvacea
  oroxylum_indicum = "data/effects_sd_map_oroxylum_indicum.R", 
  # Oroxylum indicum
  clerodendrum_quadriloculare = "data/effects_sd_map_clerodendrum_quadriloculare.R", 
  # Clerodendrum quadriloculare
  peperomia_pellucida = "data/effects_sd_map_peperomia_pellucida.R", 
  # Peperomia pellucida
  diplazium_esculentum = "data/effects_sd_map_diplazium_esculentum.R", 
  # Diplazium esculentum
  alpinia_xanthioides = "data/effects_sd_map_alpinia_xanthioides.R", 
  # Alpinia xanthioides
  curculigo_orchioides = "data/effects_sd_map_curculigo_orchioides.R", 
  # Curculigo orchioides
  ficus_racemosa = "data/effects_sd_map_ficus_racemosa.R", 
  # Ficus racemosa
  physalis_angulata = "data/effects_sd_map_physalis_angulata.R", 
  # Physalis angulata
  polygonum_perfoliatum = "data/effects_sd_map_polygonum_perfoliatum.R", 
  # Polygonum perfoliatum
  begonia_tonkinensis = "data/effects_sd_map_begonia_tonkinensis.R", 
  # Begonia tonkinensis
  begonia_aptera = "data/effects_sd_map_begonia_aptera.R", 
  # Begonia aptera
  canarium_bengalense = "data/effects_sd_map_canarium_bengalense.R", 
  # Canarium bengalense
  canarium_nigrum = "data/effects_sd_map_canarium_nigrum.R", 
  # Canarium nigrum
  canarium_tramvangum = "data/effects_sd_map_canarium_tramvangum.R", 
  # Canarium tramvangum
  ficus_heterophyllus = "data/effects_sd_map_ficus_heterophyllus.R", 
  # Ficus heterophyllus
  wedelia_trilobata = "data/effects_sd_map_wedelia_trilobata.R", 
  # Wedelia trilobata
  coix_lacryma_jobi = "data/effects_sd_map_coix_lacryma_jobi.R", 
  # Coix lacryma-jobi
  arundinaria_baviensis = "data/effects_sd_map_arundinaria_baviensis.R", 
  # Arundinaria baviensis
  ampelocalamus_patellaris = "data/effects_sd_map_ampelocalamus_patellaris.R", 
  # Ampelocalamus patellaris
  bambusa_blumeana = "data/effects_sd_map_bambusa_blumeana.R", 
  # Bambusa blumeana
  dendrocalamus_velutinus = "data/effects_sd_map_dendrocalamus_velutinus.R", 
  # Dendrocalamus velutinus
  momordica_cochinchinensis = "data/effects_sd_map_momordica_cochinchinensis.R", 
  # Momordica cochinchinensis
  clerodendrum_kaempferi = "data/effects_sd_map_clerodendrum_kaempferi.R", 
  # Clerodendrum kaempferi
  clerodendrum_philippinum = "data/effects_sd_map_clerodendrum_philippinum.R" 
  # Clerodendrum philippinum
)

```

## Create `species_data` 

As a last step we apply the `add_species_data.R` function to add the initial species data to our existing datasets (where missing, which is not usually the case). 

```{r add_species_data}
source("functions/add_species_data.R")

```

Start with an empty file

```{r start_species_data}
# Initialize an empty dataset
species_data <- data.frame()
```

Apply `add_species_data` to add default values where any are missing.

```{r loop_through_species_files}

# Loop through species and add their data
for (species in names(species_files)) {
  species_data <- add_species_data(
    existing_data = species_data,
    species_name = species,
    decision_options = decision_options,
    outcomes = outcomes,
    species_file_path = species_files[[species]]
  )
}

```

###  Load the built data and change format (for reading)

```{r load_built_data}
species_params <- as.data.frame(species_data)
#change to numeric
species_params$Effect <- as.numeric(species_params$Effect)
species_params$Variability_SD <- as.numeric(species_params$Variability_SD)
```

### Clean the species level data (update botanical names). 

```{r tolower}
species_params$Species <- tolower(trimws(species_params$Species))
```

Reverse the expenses, since higher expenses are worse, they should appear as a negative result in the plot. 

### Plot an aggregated plot with `aggregate_species_data`

```{r aggregate_plot}
source("functions/aggregate_species_data.R")
# Aggregate data across all species
aggregated_data <- aggregate_species_data(species_files)

# Summarize aggregated data
summary_data <- aggregated_data %>%
  group_by(Decision, Outcome) %>%
  summarize(
    MeanEffect = mean(Effect, na.rm = TRUE),
    MeanSD = mean(SD, na.rm = TRUE),
    .groups = "drop"
  )

```

<!-- Apply `plot_aggregate_effects` -->

<!-- ```{r plot_aggregate_effects} -->
<!-- library(ggplot2) -->
<!-- source("functions/plot_aggregate_effects.R") -->
<!-- # Plot the data -->
<!-- plot_aggregate_effects(summary_data) -->

<!-- ``` -->
