

```{r libraries_update}
library(ggplot2)
library(reshape2)
library(tidyverse)
```

# Working with data from Bavi

Collected over several visits with elders and local forest food experts, chefs, teachers, youth and other villagers. 

```{r observed_data}
# Observed data
data <- read.csv("data/observation_data_bavi.csv")
```

The data is in Vietnamese so we translate it to English. 

```{r translate_names}
# translate names(data)

data_names <- c(
  "Family.plant", "Scientific.name", "Scientific.name.Yen.Hoai.edit",
  "Local.name", "Common.name", "Local.name...Dao.language.", "Habitat",
  "Lifeform", "Uses.To.cure.what.", "Value", "Parts used", "Usage", "Frequency
  of use" , "Number of users" , "Current reserves" , "Harvest sources" ,
  "Cultivation techniques" , "Market supply potential" , "Discovery of
  processed dishes" , "Nutritional value"
)

# Convert to snake_case (replace spaces and dots with underscores, make lowercase)
cleaned_names <- gsub("[ .]+", "_", tolower(data_names))
# Assign cleaned column names directly to the data frame
names(data) <- cleaned_names
```

### Extract relevant information from the data

Look for all the words to be translated with `unique(data$current_reserves)`. Add them to the dictionary in `dictionaries.R` and translate with `translate_column`.

```{r define_dictionary}
# Source dictionary
source("functions/dictionaries.R")
# function to translate the column
source("functions/translate_column.R")
```

If `current_reserves` is `Few` or `Not many` (species is rare) we can expect positive `SpeciesConservation` effects with more restrictions. `control_access` and `provide_seedlings` will have a higher effect with lower SD, all the other options will likely be reduced. The effect would be similar for `SustainableFoodEnv`. `DietaryDiversity` would not be likely to change much. `ParkExpenditure` might also go up with more control. `TraditionalKnowledge` would likely not change much  since it is already a rare plant. If `current_reserves` are `Many` then the effects will be the opposite. 

```{r translate_current_reserves}
data$current_reserves <- translate_column(data$current_reserves, translation_dict2)
```

## Updating with parts_used data

If `parts_used` are roots, stems or whole plants then we can translate this as `destructive` and all other uses as `non_destructive`. The former type of use is more likely to have a negative `SpeciesConservation` effect with more access allowed and will shift the result.  

```{r translate_parts_used}
data$parts_used <- translate_column(data$parts_used, translation_dict3)
  # destructive_harvest <- any(data$scientific_name_snake_case == species_name & data$parts_used == "destructive")
```

If `nutritional_value` is high then the effects on `DietaryDiversity` will be raised. If `Low` effects on `DietaryDiversity` will be lowered. 

```{r translate_nutritional_value}
translate_column(data$nutritional_value, translation_dict2)

```

If `number_of_users` is `Many` then `TraditionalKnowledge` Effect would be lower and SD would be narrower for `control_access`. If it is `Few` then `DietaryDiversity` effects will be lower. 

```{r translate_number_of_users}
translate_column(data$number_of_users, translation_dict2)

```

Additional functions to clean the scientific names (these are very fiddly in the rest of the steps).

```{r clean_name_scientific_names}

  data$scientific_name <- as.character(data$scientific_name)
  
  # Function to clean and convert to snake_case
  clean_name <- function(name) {
    name <- trimws(name)  # Remove leading/trailing spaces
    name <- tolower(name)  # Convert to lowercase
    name <- gsub("[ .-]+", "_", name)  # Replace spaces and dots with underscores
    return(name)
  }
  
  # Apply function to all names
  data$scientific_name_snake_case <- sapply(data$scientific_name, clean_name, USE.NAMES = FALSE)
  
data$scientific_name_snake_case <- stringr::str_trim(data$scientific_name_snake_case)

```

# Bayesian Updating

Use `Normal-Gaussian conjugate updating` for means and standard deviations.
Here we use `tidyverse` `dplyr` and `purrr` for clean data transformations.
The `Effects` and `SDs` are adjusted based on species-specific conditions. The `update_effects_bayesian` function assumes the prior effect is normally distributed. It incorporates new information (condition-based adjustments) using a weighted update. `map()` from `purrr` updates all the expected effects from the observations in four ways: 
- High/Low Nutrition: Modifies effect and SD accordingly.
- User Count Impact: Adjusts effect size and uncertainty.
- This Bayesian approach dynamically integrates observed data into effect estimation while preserving uncertainty, ensuring a structured update.

```{r considerations_for_updating}

  # Check conditions based on the observed data

species_name = "bidens_pilosa"
rare_species <- any(data$scientific_name_snake_case == species_name &data$current_reserves == "Few")
rare_species <- any(data$scientific_name_snake_case == species_name &data$current_reserves == "Not many")
# If `current_reserves` is `Few` or `Not many` (species is rare) we can expect positive `SpeciesConservation` effects with more restrictions. `control_access` and `provide_seedlings` will have a higher effect with lower SD, all the other options will likely be reduced. The effect would be similar for `SustainableFoodEnv`. `DietaryDiversity` would not be likely to change much. `ParkExpenditure` might also go up with more control. `TraditionalKnowledge` would likely not change much  since it is already a rare plant.
common_species <- any(data$scientific_name_snake_case == species_name &data$current_reserves == "Many")
# If `current_reserves` are `Many` then the effects will be the opposite.
  destructive_harvest <- any(data$scientific_name_snake_case == species_name & data$parts_used == "destructive")
  # If `parts_used` are roots, stems or whole plants then we can translate this as `destructive` and all other uses as `non_destructive`. The former type of use is more likely to have a negative `SpeciesConservation` effect with more access allowed and will shift the result.
  high_nutrition <- any(data$scientific_name_snake_case == species_name & data$nutritional_value == "High")
  # If `nutritional_value` is high then the effects on `DietaryDiversity` will be raised.
  low_nutrition <- any(data$scientific_name_snake_case == species_name & data$nutritional_value == "Low")
  # If nutritional_value is `Low` effects on `DietaryDiversity` will be lowered.
  high_number_of_users <- any(data$scientific_name_snake_case == species_name & data$number_of_users == "Many")
  # If `number_of_users` is `Many` then `TraditionalKnowledge` Effect would be lower and SD would be narrower for `control_access`.
  few_users <- any(data$scientific_name_snake_case == species_name & data$number_of_users == "Few")
  # If number_of_users is `Few` then `DietaryDiversity` effects will be lower.
```

## Check for species consistency

Check that the species names in the prior data `species_files` and the new cleaned species names in the `data` column `scientific_name_snake_case` have the same names. Check which are different and justify the missing from each list. 

```{r check_missing_priors, echo = FALSE, message=TRUE}
# Extract species names from original prior data species_files list
species_files_list <- names(species_files)

# Extract unique species names from the data from the field
species_data_list <- unique(data$scientific_name_snake_case)
```

### Species not included in first rounds

```{r present_in_prior_not_obs}
# Find species missing in prior species_files but present in observation data
setdiff(species_data_list, species_files_list)
```

Missing species in `species_files` (present in `data` but not in `species_files`) are plants that are common but not really relevant for the question of access to forest species and so are not included in the assessment. 

### Species not included in later rounds

```{r present_obs_not_prior}
# Find extra species in prior species_files but not in observation data
setdiff(species_files_list, species_data_list)
```

Extra species in `species_files` (present in `species_files` but not in `data`) are plants that we recognized a food use for early on in field explorations but did not learn any more about them in later visits. 