---
title: "Forest access for sustainable food environments in Bavi"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
library(ggplot2)
library(reshape2)
library(tidyverse)
```

Here we use Monte Carlo simulations based on expert knowledge. This is a well-established and flexible method for decision modeling. Allowing for use of expert-derived probabilities and variances to estimate outcomes, even when data is sparse. With this method we can capture the variability and uncertainty in model inputs. We can apply this to guide decision-making by identifying likely outcomes and their variability.

# Generalized Function 

```{r child="A_Generalized_function.Rmd"}
```

# Prior Estimates 

```{r child = "B_Prior_estimates.Rmd"}
```

# Decision model {#decision_model}

```{r child = "C_Decision_model.Rmd"}
```

# Update based on field observations {#update_model}

```{r child = "D_Update.Rmd"}
```

# Generalize the update to all species in the collection (from all trips)

```{r child = "E_Update_all_species.Rmd"}
```
